Report

### 2022-05-13 manuscript.Rmd
update 2022-05-13 20:25:20

```{r setup, include = FALSE}
run_everything <- FALSE
library("papaja")
library(tidySEM)
library(kableExtra)
r_refs("r-references.bib")
```

@@

The present simulation study set out to validate the PBF algorithm.
For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
This correlation coefficient was either equal to or larger than an informative hypothesis value.
In this study, the informative hypothesis was kept constant, and specified as $H_i: \rho > .1$.
We then computed a PBF to determine the amount of evidence in favor of $H_i$ across these multiple samples.
A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was a naive "vote count" algorithm,
which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
The second was a random-effects meta-analysis (RMA), which is the standard in the field.
For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
Third was an individual participant data (IPD) meta-analysis.
Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
In contrast to classic meta-analysis, however, IPD freely estimates variance at the first level, because raw data are available.
The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
It is thus informative to compare the PBF to both these methods.
IPD was also evaluated using a 90% confidence interval for the overall effect size.

@@

To examine performance in a range of realistic scenarios,
several design factors were manipulated:
The population effect size $\rho \in (-0.1, 0, 0.1, 0.2)$,
the sample size per group $n \in (20, 80, 200, 500)$,
the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is the assumed when analyzing correlations between observed items or scale scores.
The design factors combined to produce `r out$conditions` unique conditions.
For all simulation conditions, 1000 data sets were generated.

```{r, echo = F}
library(data.table) # load in libraries
library(DT)
library(ggplot2)
rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")

@@

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
First, we examined overall model performance across conditions.
These results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
In contrast, specificity was very high for all algorithms except PBF.
This suggests that the other algorithms classified most conditions as negatives (no effect found),
regardless of the existence of a population effect.
The PBF trades a loss of specificity for increased sensitivity,
and average levels of both were approximately equal.
As the PBF vastly outperformed other BFs, these were omitted from further analysis.

@@

We examined the effect of simulation conditions on overall accuracy.

```{r tabeffect, echo = FALSE}
tab <- read.csv("effect_of_conditions.csv", stringsAsFactors = FALSE)
tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
tab <- tab[, c("condition", "prodbf_ic", "rma", "ipd", "allsig", "prodbf_ic.vs..rma", "ipd.vs..prodbf_ic", "allsig.vs..prodbf_ic")]
names(tab) <-  c("Factor", "PBF", "RMA", "IPD", "VC", "vs RMA", "vs IPD", "vs VC")

knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.

@@

```{r fign, fig.cap="Mean performance by sample size"}
res <- readRDS("confusion_by_cond.RData")
res <- res[res$alg %in% c("ipd", "rma", "allsig", "prodbf_ic"), ]

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Sample size")
```

@@

```{r, fig.cap="Mean performance by number of groups"}
df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Number of groups")
```
The figure indicates that, for PBF at higher levels of k, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
This difference in pattern of effects explains why number of groups has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC shows decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining any false negatives increases with the number of groups.

@@

```{r, fig.cap="Mean performance by reliability"}

# dat <- readRDS(file.path("Sim", "sim_results_2022-05-07.RData"))
# varsout <- c("ipd", "rma", "allsig", "prodbf_ic")
# varspred <- c("errorsd", "n", "k", "hyp_val")
# 
# df_plot <- dat[!es < .1,  lapply(.SD, function(var) {
#   as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
# }), .SDcols = varsout, by = "errorsd"]
# 
# tabres[, metric := rep(c("TN", "FN", "FP", "TP"), nrow(tabres)/4)]
# 
# tabres <- melt(tabres, measure.vars = varsout,
#              variable.name = "alg", value.name = "val")
# tabres <- dcast(tabres, errorsd + n + k + hyp_val + alg ~ metric, value.var = "val")
# tabres[, alpha := FP / (FP + TN)]
# tabres[, beta := FN / (TP + FN)]
# tabres[, sensitivity := 1-beta]
# tabres[, specificity := 1-alpha]
# tabres[, accuracy := (TP + TN) / (TP+TN+FP+FN)]

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("errorsd", "Outcome", "alg")]
ggplot(df_plot, aes(x = errorsd, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Reliability")
```

>>>>

```{r setup, include = FALSE}
run_everything <- FALSE
library("papaja")
library(tidySEM)
library(kableExtra)
r_refs("r-references.bib")
out <- readRDS("out.RData")
```

@@

The present simulation study set out to validate the PBF algorithm.
For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
The informative hypothesis was kept constant at $H_i: \rho > .1$.
All simulation conditions were evaluated once in the presence of a true population effect,
defined as $\rho = .2$,
and once in the presence of a null effect, defined as $\rho = .1$.
A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was a "vote count" algorithm,
which is common but not considered to be good practice [REF].
This approach used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
The second was a random-effects meta-analysis (RMA), which is the standard in the field.
For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
Third was an individual participant data (IPD) meta-analysis.
Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
In contrast to classic meta-analysis, however, IPD freely estimates variance at the first level, because raw data are available.
The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
It is thus informative to compare the PBF to both these methods.
IPD was also evaluated using a 90% confidence interval for the overall effect size.

@@

To examine performance in a range of realistic scenarios,
several design factors were manipulated:
The sample size per group $n \in (20, 80, 200, 500)$,
the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is the assumed when analyzing correlations between observed items or scale scores.
The design factors combined to produce `r out$simreps/1000` unique conditions.
For all simulation conditions, 1000 data sets were generated.

```{r, echo = F}
library(data.table) # load in libraries
library(DT)
library(ggplot2)
# rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")

@@

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
# names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
First, we examined overall model performance across conditions.
These results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
In contrast, specificity was very high for all algorithms except PBF.
This suggests that the other algorithms classified most conditions as negatives (no effect found),
regardless of the existence of a population effect.
The PBF trades a loss of specificity for increased sensitivity,
and average levels of both were approximately equal.
As the PBF vastly outperformed other BFs, these were omitted from further analysis.

@@

We examined the effect of simulation conditions on overall accuracy.
```{r tabeffect, echo = FALSE}
tab <- read.csv("effect_of_conditions.csv", stringsAsFactors = FALSE)
tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
tab <- tab[, grepl("^\\w+$", names(tab)) | (grepl("\\.{2}", names(tab)) & grepl("PBF", names(tab)))]
names(tab)[grepl("\\.{2}", names(tab))] <- paste0("vs ", gsub("(\\.|PBF|vs)", "", names(tab)[grepl("\\.{2}", names(tab))]))
knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.

@@

```{r fign, fig.cap="Mean performance by sample size"}
res <- readRDS("confusion_by_cond.RData")

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("n", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Sample size")
```

@@

```{r, fig.cap="Mean performance by number of groups"}
df_plot <- data.table::melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("k", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Number of groups")
```
The figure indicates that, for PBF at higher levels of k, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
This difference in pattern of effects explains why number of groups has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC shows decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining any false negatives increases with the number of groups.

@@

```{r, fig.cap="Mean performance by reliability"}
df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("reliability", "Outcome", "alg")]
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("reliability", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
ggplot(df_plot, aes(x = reliability, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Reliability")
```
